---
title: "COVID-19 in Saudi Arabia"
description: |
  The coronavirus hit the Kingdom of Saudi Arabia
  When was the first cases? How it affect the Middle East country?
author:
  - name: Putu Wahyu Saputra 
    email: psap0002@student.monash.edu
date: 09-14-2020
output:
  distill::distill_article:
    self_contained: false
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, options(scipen=999))
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lubridate)
library(coronavirus) # update_dataset()
library(htmltools)
library(plotly)
library(tidycovid19)
library(scales)
```


```{r, include=FALSE}
#load coronavirus dataset
data("coronavirus")
saudi <- coronavirus %>% 
  filter(country=="Saudi Arabia", date <= "2020-09-10") %>%
  select(-province)

saudi_test_per1000 <- read_csv("daily-tests-per-thousand-people-smoothed-7-day.csv") %>% filter(Code=="SAU")
saudi_test <- download_owid_testing_data() %>% filter(iso3c == "SAU")

```

```{r,include = FALSE}
variables_1 <- saudi %>% names()
class_1 <- sapply(saudi, class)
desc_1 <- c("The date of the summary","The province or state, when applicable","Latitude point","Longitude point","the type of case (confirmed, recovered, death)","the number of daily cases (corresponding to the case type)")

variables_2 <- saudi_test %>% names()
class_2 <- sapply(saudi_test, class)
desc_2 <- c("Country code","The date of the summary","	Accumulated test counts", "Definition of what constitutes a ‘test’", "The date and time of the summary" )
```

## Dataset

This `coronavirus` dataset has the following variables as seen in Table \@ref(tab:corona-tab-vr). The dataset is obtained from `coronavirus` package in CRAN. @corona

```{r corona-tab-vr, echo=FALSE}
tibble(variables_1,class_1,desc_1) %>%
  kable(col.names = c("variable","class","description"),
        caption ="Coronavirus dataset variables") %>%
  kable_styling(bootstrap_options = "hover")
```

To make comparisons, the package from `tidycovid19` is also used in this report. @tidycovid19 This package has more variables because it is a combination of several other dataset sources, including the Our World in Data. The dataset contains a tidy format of the testing data over time. The variables from Our World in Data listed below in Table \@ref(tab:tidycovid-tbl).

```{r tidycovid-tbl}
tibble(variables_2,class_2,desc_2) %>%
  kable(col.names = c("variable","class","description"),
        caption ="Our World dataset variables") %>%
  kable_styling(bootstrap_options = "hover")
```


## When was the first cases and how the trend overtime ?

AA reported, the first case of coronavirus in Saudi Arabia occurred on March 2, 2020, where a citizen was allegedly coming from Iran via Bahrain. @anodolu In the following three days, the second case was reported as many as 4 cases at once. In less than two weeks since the first reported case, the total number of cases has increased by almost six times higher. This trend continued to increase daily, reaching a peak on May 16 with daily cases of 2870. After that cases started to decline for almost two weeks recording 1581 cases a day on May 29. Cases have risen sharply since then, hitting a record high of 4919 cases per day on June 17. After reaching the peak, the curve trend began to decline, although it fluctuated at the beginning of July. From the latest data collected on September 9, cases per day have reached 708 cases per day.

We provide a plot showing the fatality trend in Saudi Arabia in Figure xx. The first reported death cases caused by the coronavirus occurred 20 days after the first case was reported where the victim was a 51-year-old Afghan resident @aljazeera Trend fatality increased after that to reach a peak of 58 deaths a day on July 5. After that, the trend of mortality began to decline even though it was fluctuating.

Apart from positive cases and deaths, Saudi Arabia is showing an increasing trend in patient recovery. The trend increased from the end of April and had decreased in June. Patient recovery per day tends to increase despite the fluctuation and peaked on July 14. On that date, 7718 people were recorded as recovered from the coronavirus. As the number of positive cases decreased, the number of recovered patients also decreased.AA reported, the first case of coronavirus in Saudi Arabia occurred on March 2, 2020, where a citizen was allegedly coming from Iran via Bahrain. @anodolu In the following three days, the second case was reported as many as 4 cases at once. In less than two weeks since the first reported case, the total number of cases has increased by almost six times higher. This trend continued to increase daily, reaching a peak on May 16 with daily cases of 2870. After that cases started to decline for almost two weeks recording 1581 cases a day on May 29. Cases have risen sharply since then, hitting a record high of 4919 cases per day on June 17. After reaching the peak, the curve trend began to decline, although it fluctuated at the beginning of July. From the latest data collected on September 9, cases per day have reached 708 cases per day.

We provide a plot showing the fatality trend in Saudi Arabia in Figure xx. The first reported death cases caused by the coronavirus occurred 20 days after the first case was reported where the victim was a 51-year-old Afghan resident @aljazeera Trend fatality increased after that to reach a peak of 58 deaths a day on July 5. After that, the trend of mortality began to decline even though it was fluctuating.

Apart from positive cases and deaths, Saudi Arabia is showing an increasing trend in patient recovery. The trend increased from the end of April and had decreased in June. Patient recovery per day tends to increase despite the fluctuation and peaked on July 14. On that date, 7718 people were recorded as recovered from the coronavirus. As the number of positive cases decreased, the number of recovered patients also decreased.


```{r daily-cases, fig.cap="Daily cases in Iraq from January until August, overall the trend of cases is raising until reached the peak in June after that it declined overtime.", echo=FALSE}
plot <- saudi %>%
  ggplot(aes(x=date,
             y=cases,
             color=type)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  ggtitle("Daily cases in Saudi Arabia")  +
  scale_x_date(date_breaks = "1 month", 
                 labels=date_format("%b-%Y"),
                 limits = as.Date(c('2020-05-13','2020-09-13')))


  ggplotly(plot) %>%
    layout(legend = list(orientation = "h", x = 0.3, y = -0.2)) %>%
    config(displayModeBar = FALSE)
```

```{r death-saudi, fig.cap="The daily deaths in Saudi increased from April and reached a peak in July, after which case deaths began to decline through September despite the volatility"}
death <- saudi %>%
  filter(type=="death") %>%
  ggplot(aes(x=date,
             y=cases,
             color=type)) +
  geom_line() +
  theme_minimal() +
    theme(legend.position = "none") +
  ggtitle("Fatality in Saudi Arabia") +
  scale_x_date(date_breaks = "1 month", 
                 labels=date_format("%b-%Y"),
                 limits = as.Date(c('2020-05-13','2020-09-13')))

ggplotly(death)%>%
    layout(legend = list(orientation = "h", x = 0.3, y = -0.2)) %>%
    config(displayModeBar = FALSE)
```

```{r summary}
summary <- saudi %>%
  pivot_wider(names_from = "type", values_from="cases") %>%
  summarise(total_confirmed = sum(confirmed),
            total_death= sum(death),
            total_recovered=sum(recovered)) %>%
  mutate(active_cases = total_confirmed-total_death-total_recovered)
summary %>%
kable(caption = "The summary of COVID-19 cases in Saudi Arabia in 9 months, left with 19,533 active cases in the whole country", col.names = c("Total Confirmed", "Total Death", "Total Recovered", "Active Cases")) %>%
  kable_styling("hover")
```

Until this data is recorded on September 10, in Table \@ref(tab:summary) there are total `r summary$total_confirmed` positive cases,` r summary$total_death` deaths, `r summary$total_recovered` recovered, and` r summary$active_cases` active case.

## Test Performed

Given the relatively high cure rate with a low fatality, let us see how well Saudi Arabia has responded to this pandemic by looking at the number of tests it has conducted. Doing tests to identify the virus in a person is very important to help to trace clusters. Test data was performed downloaded from the Our World in Data website. @owidcoronavirus This data contains the total number of tests performed. Figure \@ref(fig:test-1000) shows the daily COVID-19 test per thousand people (The figures are given as a rolling 7-day average.)

In general, the test performed has an increasing trend. The trend tests carried out increased significantly from May to July. After that, the trend experienced a period of fluctuation and increased again and reached a peak on August 16, with 1,880 daily tests per thousand people. As cases decreased, the test performed also decreased. Until the latest data on September 9, an estimate of 1.3 daily tests per thousand people has been carried out.

```{r test,  fig.show="hold", out.width="50%"}
plot_test1000 <- saudi_test_per1000 %>%
  ggplot(aes(x=Date,
             y=new_tests_per_thousand_7day_smoothed))+
  geom_point(color="blue")+
  geom_line()+
  theme_minimal()+
  ylab("Test per 1000 People") +
  xlab("") +
  labs(tags="(A)", title = "Daily Tests per 1000 People") +
  scale_x_date(date_breaks = "1 month", 
                 labels=date_format("%b-%Y"),
                 limits = as.Date(c('2020-05-13','2020-09-10'))) +
  theme(plot.tag = element_text(size=11),
    plot.title = element_text(size=11, face = "bold"),
    axis.title.y = element_text(size=10))

plot_total_test <- saudi_test %>%
  ggplot(aes(x=date,
             y=total_tests)) +
  geom_col(width = 0.259,color="blue")+
  theme_minimal() +
  ylab("Total Test")+
  xlab("Date") +
  theme(legend.title = element_blank(),
        plot.tag = element_text(size=11),
    plot.title = element_text(size=11, face = "bold"),
    axis.title.y = element_text(size=10))+
  labs(tags="(B)", title="Total Cummulative Test")+
  scale_x_date(date_breaks = "1 month", 
                 labels=date_format("%b-%Y"),
                 limits = as.Date(c('2020-05-13','2020-09-13')))
```

```{r}
library(patchwork)
plot_test1000 + plot_total_test + plot_layout(nrow=2, guides = "collect")
```


## How it affected Saudi?

```{r}
hajj <- read_csv2("domestic-and-foreign-pilgrims.csv")
hajj %>% filter(Year>2010) %>%
  ggplot(aes(x=Year,
         y=`Number of Pilgrims`,
         fill=`Domestic/Foreign Pilgrim`))+
  geom_col(position = "dodge")
```




# Acknowledgements

This blog post constructed by R package as following:
- distill @distill
- tidyverse @tidyverse2
- ggplot2 @ggplot2
- knitr @knitr
- kableExtra @kableExtra
- lubridate @lubridate
- DT @DT
- htmltools @htmltools
- plotly @plotly
- gt @gt

